---
title: "Testing Script for DGCA Parallel Functions"
author: "Max Segnitz"
date: "2024-05-29"
output: html_document
---

```{r}
# Load Customization of DGCA with improved speed & parallel handling.
source("https://raw.githubusercontent.com/rmsegnitz/Bioinformatics_Tools/master/R_functions/DGCA_parFuncs.R")

library(tidyverse)
library(kimma)
```


```{r}
# load sample data
sample_dat<-kimma::example.voom

# define simple contrasts
View(sample_dat$targets)

# define reduced module set
genes_clust<-
  sample_dat$E[heatmap(sample_dat$E)$rowInd, ]

dummy_mods<-
  data.frame(geneName=rownames(genes_clust), 
             module=c(rep("dummy_100", 100), rep("dummy_150", 150), rep("dummy_250", 250),
                      rep("dummy_500", 500)))%>%
  left_join(dplyr::select(sample_dat$genes, geneName, hgnc_symbol))

dummy_mods_DGCA<-
  dplyr::select(dummy_mods, geneName, module)%>%
  dplyr::rename(values=geneName, ind=module)%>%
  arrange(ind)
```

```{r}


#Define design matrices as in limma
design_mat <-
  model.matrix(~ 0 + virus, 
               data=sample_dat$targets)

# rename/simplify matrix column names
colnames(design_mat)<-str_remove_all(colnames(design_mat), "virus")
contrast_test<-c("none", "HRV")

system.time(
modules_DGCA_test<-
                moduleDC_par(
                  inputMat = sample_dat$E, # reduce input matrix to relevant libraries
                        design = design_mat, # reduce to relevant libraries and groups
                        compare = contrast_test,
                        genes = dummy_mods_DGCA$values, 
                        labels = dummy_mods_DGCA$ind, 
                        nPerms=1000, number_DC_genes=0,
                        dCorAvgMethod="median", corrType="pearson", 
                        num_cores=10, 
                        save_gene_level=T, out.dir="~/Desktop/DGCA_test/test_together", seed=32)
)

```

# Test Running with individualized modules

It would appear that there may be residual code in the DGCA functions that calculates ALL pairwise correlations from the entire expression matrix it receives (as opposed to only running within module). This may be intentional, and perhaps it would be unwise to skirt this, but I would like to test whether it impacts the resulting stats and run times. 

```{r}
modules_DGCA_test_perMod<-list()
system.time(
for(i in unique(dummy_mods_DGCA$ind)){
    
  mod_sub=filter(dummy_mods_DGCA, ind==i)
  expression_sub<-sample_dat$E[mod_sub$values, ]
   
modules_DGCA_test_perMod[[i]]<-
                moduleDC_par(
                  inputMat = expression_sub, # reduce input matrix to relevant libraries
                        design = design_mat, # reduce to relevant libraries and groups
                        compare = contrast_test,
                        genes = mod_sub$values, 
                        labels = mod_sub$ind, 
                        nPerms=1000, number_DC_genes=0,
                        dCorAvgMethod="median", corrType="pearson", 
                        num_cores=10, 
                        save_gene_level=T, out.dir="~/Desktop/DGCA_test/test_loop", seed=32)}
)

```

```{r}
View(modules_DGCA_test)
View(modules_DGCA_test_perMod$dummy_100)

```

